Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-02-12 PAPER.tex
moved adrian up on author list
quoted matt's best limit

\author{
Carina Cheng\altaffilmark{1}$^{,\diamond}$,
Aaron R. Parsons\altaffilmark{1,2},
Matthew Kolopanis \altaffilmark{3},
Daniel C. Jacobs\altaffilmark{3}, 
Saul A. Kohn\altaffilmark{4},
Adrian Liu\altaffilmark{1,5}$^{,\dagger}$, 
Jonathan C. Pober\altaffilmark{6}, 
James E.~Aguirre\altaffilmark{4},
Zaki S. Ali\altaffilmark{1}, 
Gianni Bernardi\altaffilmark{7,8,9}, 
Richard F. Bradley\altaffilmark{10,11,12},
Chris L. Carilli\altaffilmark{13,14},
David R. DeBoer\altaffilmark{2}, 
Matthew R. Dexter\altaffilmark{2},
Joshua S. Dillon\altaffilmark{1}$^{,*}$,
Pat Klima\altaffilmark{11},
David H. E. MacMahon\altaffilmark{2},
David F. Moore\altaffilmark{4},
Chuneeta D. Nunhokee\altaffilmark{8},
William P. Walbrugh\altaffilmark{7},
Andre Walker\altaffilmark{7}
}

\altaffiltext{1}{Astronomy Dept., U. California, Berkeley, CA}
\altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley CA}
\altaffiltext{3}{School of Earth and Space Exploration, Arizona State U., Tempe AZ}
\altaffiltext{4}{Dept. of Physics and Astronomy, U. Penn., Philadelphia PA} 
\altaffiltext{5}{Berkeley Center for Cosmological Physics, Berkeley, CA} 
\altaffiltext{6}{Dept. of Physics, Brown University, Providence RI}
\altaffiltext{7}{Square Kilometer Array, S. Africa, Cape Town South Africa}
\altaffiltext{8}{Dept. of Physics and Electronics, Rhodes University, South Africa}
\altaffiltext{9}{INAF-Instituto di Radioastronomia, Bologna Italy}
\altaffiltext{10}{Dept. of Electrical and Computer Engineering, U. Virginia, Charlottesville VA}
\altaffiltext{11}{National Radio Astronomy Obs., Charlottesville VA}
\altaffiltext{12}{Dept. of Astronomy, U. Virginia, Charlottesville VA}
\altaffiltext{13}{National Radio Astronomy Obs., Socorro NM}
\altaffiltext{14}{Cavendish Lab., Cambridge UK}

@@

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the first stars and 
galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the 
EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation 
of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to 
develop techniques that help maximize sensitivity while validating results. Additionally, it is imperative to understand the trade-offs 
between different methods and their effects on power spectrum estimates. In this paper, we focus on three major 
themes --- signal loss, power spectrum error estimation, and bias in measurements. We describe techniques that affect 
these themes using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array 
for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have 
led to a revised, higher $21$ cm power spectrum upper limit from PAPER-64. This revised result, presented in a companion paper by Kolopanis et al., \textit{in prep}, mostly stems from an 
improved signal loss calculation for loss associated with empirically estimated covariances and supersedes results from 
previously published PAPER analyses.
\end{abstract}

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, errors, and bias) that are essential to investigate 
for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques (fringe-rate filtering, weighting, bootstrapping, jackknife testing) and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned. More specifically, we develop new techniques for computing signal loss and under-go a rigorous re-analysis of PAPER-64 data in order to present a new power spectrum limit. A companion paper, Kolopanis et al., \textit{in prep}, builds off of the methods in this paper to present PAPER-64 results for multi-redshifts and multiple baselines.

@@

This paper is organized into two main sections. In Section \ref{sec:Themes} we introduce the three themes of our focus, using a 
toy model to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data 
from the PAPER-64 array, highlighting key changes from the previously published result in \citet{ali_et_al2015}, henceforth known as \citetalias{ali_et_al2015}, which have led to a 
revised PAPER-64 power spectrum result (Kolopanis et al., \textit{in prep}). We conclude in Section \ref{sec:Con}.

@@

\noindent of $(22.4$ mK$)^{2}$ in the range $0.15 < k < 0.5$ h Mpc$^{-1}$ at $z = 8.4$. The revision of this limit (Kolopanis et al., \textit{in prep}) stems from previously underestimated signal loss and underestimated error bars, both of which we 
address in the following sections.

@@

Our revised, best to-date PAPER-64 power spectrum (using only one baseline separation type and $\widehat{\textbf{C}}_{eff}$) is shown in Figure 
\ref{fig:ps1_data}. Again, black and gray points correspond to positive and negative power spectrum values respectively, with 
$2\sigma$ errors bars. Also plotted are the uniform-weighted power spectrum upper limit (dashed blue) and theoretical prediction of 
noise (solid green). From this result, we quote a best $2\sigma$ upper limit of $(151.0$ mK$)^{2}$ at $k=0.34$ $h$Mpc$^{-1}$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim7$ in mK (though we only use one baseline type in our analysis as opposed to three). 
\cc{replace with kolopanis limit instead later}

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.png}
	\caption{Power spectrum of PAPER-64 using $\widehat{\textbf{C}}_{eff}$. Black and gray points correspond to positive and 
negative power spectrum values, respectively, with $2\sigma$ error bars also plotted. The dashed blue line is the uniform-weighted 
power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on 
observational parameters. This power spectrum result differs from \citetalias{ali_et_al2015} in that it only uses data from one type of baseline ($30$ m 
East/West baselines) instead of three. Major differences from previously published results stem from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. From this, we quote a best $2\sigma$ upper limit of $(151.0$ mK$)^{2}$ at 
$k=0.34$ $h$Mpc$^{-1}$, a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim7$ in mK \cc{replace with kolopanis limit later}.}
	\label{fig:ps1_data}
\end{figure*}

In this section we have shown three simple ways of regularizing $\widehat{\textbf{C}}$ to minimize signal loss using PAPER-64 
data. There are many other weighting schemes that we leave for consideration in future work. For example, one could estimate 
$\widehat{\textbf{C}}$ using information from different subsets of baselines. For redundant arrays this could mean calculating $
\widehat{\textbf{C}}$ from a different but similar baseline type, such as the $\sim30$ m diagonal PAPER baselines (instead of the 
horizontal E/W ones). Alternately, covariances could be estimated from all other baselines except the two being cross-multiplied 
when forming a power spectrum estimate. This method was used in \citet{parsons_et_al2014} (a similar method was also used in \citet{dillon_et_al2015}) in order to avoid suppressing the 
$21$ cm signal, and it's worth noting that the PAPER-32 results are likely less impacted from the issue of signal loss underestimation 
because of this very reason (however, they are affected by the error estimation issues described in Section \ref{sec:Error}, so 
we also regard those results as suspect and superseded by those of Kolopanis et al., \textit{in prep}).

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $144$ K, yielding $T_{sys} = 431$ K at $150$ MHz. 
This is lower than the $T_{sys}$ of $500$ K used in \citetalias{ali_et_al2015} because of several small miscalculation errors that were 
identified\footnote{For example, there was a missing a square root in going from a variance to a standard deviation.}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after cross-multiplying independent `even' and `odd' visibility measurements (which is done principally to avoid a noise bias). In \citetalias{ali_et_al2015}, a factor of $2$ was mistakenly used.
\item $N_{lst}$: The number of LST bins that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{lst} = 8$.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The updated limits in Kolopanis et al., \textit{in prep}
use three separation types.
\item $t_{int}$: Length of an independent integration of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time $t_{i}$
using the following:
\begin{equation}
t_{int} = t_{i} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{i}=43$ seconds, $t_{int}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{int} = 3857$ s. 
\item $N_{days}$: The total number of days of data analyzed. In \citetalias{ali_et_al2015}, this number was set to $135$. However, because we 
divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, 
because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a more realistic value of 
$N_{days}$ as:
\begin{equation}
 \frac{1}{N_{days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ is over LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ 
days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In \citetalias{ali_et_al2015}, this number was 
the total number of $30$ m East/West baselines used in the analysis. However, using the total number of baselines neglects 
the fact that baselines are averaged into groups (see Section \ref{sec:Boot}) for computational speed-up when cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{bls} = \frac{N_{bls\_total}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}},
\end{equation}
\noindent where, for the \citetalias{ali_et_al2015} analysis, $N_{gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{bls\_total}/N_{gps}$) and then as the square root of the number of cross-multiplied pairs ($\sqrt{N_{gps}^{2} - 
N_{gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$ instead 
of $51$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}

@@

As a consequence of our investigations, we have also revised power spectrum results from PAPER-64. For an analysis of only 
one baseline type, we quote a new $2\sigma$ upper limit of $(151.0$ mK$)^{2}$ at $k=0.34$ $h$Mpc$^{-1}$, a higher limit than 
\citetalias{ali_et_al2015} by a factor of $\sim7$ in mK \cc{replace with kolopanis limit later}. The reasons for a previously underestimated limit and 
ways in which our new analysis differs can be summarized by the following:

\begin{itemize}
\item Signal loss, previously found to be $<2\%$ in \citetalias{ali_et_al2015}, was under-estimated by a factor of $\sim$$1000$ for inverse 
covariance weighting. For our new analysis, we use a regularized covariance weighting method to minimize loss (Section 
\ref{sec:Weight}). However, because our revised weighting method is not as aggressive as the former, our results are still a 
factor of $\sim7$ in mK higher than previous limits. Under-estimated signal loss therefore represents the bulk of our revision. We 
note that PAPER's analysis is not the first to under-estimate loss; results from the GMRT (\citealt{paciga_et_al2013}) were also 
revised from new signal loss calculations associated with their singular value decomposition foreground filter.
\item Power spectrum errors, originally computed by bootstrapping, were under-estimated for the data by a factor of $\sim$$2$ in mK due to oversampling data whose effective number of independent samples was reduced from fringe-rate filtering (Section \ref{sec:Boot}). 
\item Several factors used in an analytic expression to predict the noise-level in PAPER-64 data were revised, yielding a 
decrease in predicted sensitivity level by a factor of $\sim$$2$ in mK (Section \ref{sec:PSSense}). We have verified our revised 
prediction extensively using pure noise simulations. We note that our sensitivity prediction is revised by a factor less than our 
power spectrum result, implying that if taken at face value, the theoretical prediction for noise in \citetalias{ali_et_al2015} was too high for its data 
points.
\end{itemize}

>>>>

\author{
Carina Cheng\altaffilmark{1}$^{,\diamond}$,
Aaron R. Parsons\altaffilmark{1,2},
Matthew Kolopanis \altaffilmark{3},
Daniel C. Jacobs\altaffilmark{3}, 
Adrian Liu\altaffilmark{1,4}$^{,\dagger}$, 
Saul A. Kohn\altaffilmark{5},
Jonathan C. Pober\altaffilmark{6}, 
James E.~Aguirre\altaffilmark{5},
Zaki S. Ali\altaffilmark{1}, 
Gianni Bernardi\altaffilmark{7,8,9}, 
Richard F. Bradley\altaffilmark{10,11,12},
Chris L. Carilli\altaffilmark{13,14},
David R. DeBoer\altaffilmark{2}, 
Matthew R. Dexter\altaffilmark{2},
Joshua S. Dillon\altaffilmark{1}$^{,*}$,
Pat Klima\altaffilmark{11},
David H. E. MacMahon\altaffilmark{2},
David F. Moore\altaffilmark{5},
Chuneeta D. Nunhokee\altaffilmark{8},
William P. Walbrugh\altaffilmark{7},
Andre Walker\altaffilmark{7}
}

\altaffiltext{1}{Astronomy Dept., U. California, Berkeley, CA}
\altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley CA}
\altaffiltext{3}{School of Earth and Space Exploration, Arizona State U., Tempe AZ}
\altaffiltext{4}{Berkeley Center for Cosmological Physics, Berkeley, CA} 
\altaffiltext{5}{Dept. of Physics and Astronomy, U. Penn., Philadelphia PA} 
\altaffiltext{6}{Dept. of Physics, Brown University, Providence RI}
\altaffiltext{7}{Square Kilometer Array, S. Africa, Cape Town South Africa}
\altaffiltext{8}{Dept. of Physics and Electronics, Rhodes University, South Africa}
\altaffiltext{9}{INAF-Instituto di Radioastronomia, Bologna Italy}
\altaffiltext{10}{Dept. of Electrical and Computer Engineering, U. Virginia, Charlottesville VA}
\altaffiltext{11}{National Radio Astronomy Obs., Charlottesville VA}
\altaffiltext{12}{Dept. of Astronomy, U. Virginia, Charlottesville VA}
\altaffiltext{13}{National Radio Astronomy Obs., Socorro NM}
\altaffiltext{14}{Cavendish Lab., Cambridge UK}

@@

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the first stars and 
galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the 
EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation 
of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to 
develop techniques that help maximize sensitivity while validating results. Additionally, it is imperative to understand the trade-offs 
between different methods and their effects on power spectrum estimates. In this paper, we focus on three major 
themes --- signal loss, power spectrum error estimation, and bias in measurements. We describe techniques that affect 
these themes using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array 
for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have 
led to a revised, higher $21$ cm power spectrum upper limit from PAPER-64. This revised result, presented in a companion paper by Kolopanis et al., \textit{submitted}, mostly stems from an 
improved signal loss calculation for loss associated with empirically estimated covariances and supersedes results from 
previously published PAPER analyses.
\end{abstract}

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, errors, and bias) that are essential to investigate 
for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques (fringe-rate filtering, weighting, bootstrapping, jackknife testing) and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned. More specifically, we develop new techniques for computing signal loss and under-go a rigorous re-analysis of PAPER-64 data in order to present a new power spectrum limit. A companion paper, Kolopanis et al., \textit{submitted}, builds off of the methods in this paper to present PAPER-64 results for multi-redshifts and multiple baselines.

@@

This paper is organized into two main sections. In Section \ref{sec:Themes} we introduce the three themes of our focus, using a 
toy model to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data 
from the PAPER-64 array, highlighting key changes from the previously published result in \citet{ali_et_al2015}, henceforth known as \citetalias{ali_et_al2015}, which have led to a 
revised PAPER-64 power spectrum result (Kolopanis et al., \textit{submitted}). We conclude in Section \ref{sec:Con}.

@@

\noindent of $(22.4$ mK$)^{2}$ in the range $0.15 < k < 0.5$ h Mpc$^{-1}$ at $z = 8.4$. The revision of this limit (Kolopanis et al., \textit{submitted}) stems from previously underestimated signal loss and underestimated error bars, both of which we 
address in the following sections.

@@

A revised PAPER-64 power spectrum (using only one baseline separation type and $\widehat{\textbf{C}}_{eff}$) is shown in Figure 
\ref{fig:ps1_data}. Again, black and gray points correspond to positive and negative power spectrum values respectively, with 
$2\sigma$ errors bars. Also plotted are the uniform-weighted power spectrum upper limit (dashed blue) and theoretical prediction of 
noise (solid green). In Kolopanis et al., (\textit{submitted}), a larger parameter space is explored for regularization and multiple baselines are used in the analysis to produce a best $2\sigma$ power spectrum upper limit of $(218.7$ mK$)^{2}$ at $k=0.3$ $h$Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK.

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.png}
	\caption{Power spectrum of PAPER-64 using $\widehat{\textbf{C}}_{eff}$. Black and gray points correspond to positive and 
negative power spectrum values, respectively, with $2\sigma$ error bars also plotted. The dashed blue line is the uniform-weighted 
power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on 
observational parameters. This power spectrum result differs from \citetalias{ali_et_al2015} in that it only uses data from one type of baseline ($30$ m 
East/West baselines) instead of three. Major differences from previously published results stem from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. In Kolopanis et al., (\textit{submitted}), a larger parameter space is explored for regularization and multiple baselines are used in the analysis to produce a best $2\sigma$ power spectrum upper limit of $(218.7$ mK$)^{2}$ at $k=0.3$ $h$Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK.}
	\label{fig:ps1_data}
\end{figure*}

In this section we have shown three simple ways of regularizing $\widehat{\textbf{C}}$ to minimize signal loss using PAPER-64 
data. There are many other weighting schemes that we leave for consideration in future work. For example, one could estimate 
$\widehat{\textbf{C}}$ using information from different subsets of baselines. For redundant arrays this could mean calculating $
\widehat{\textbf{C}}$ from a different but similar baseline type, such as the $\sim30$ m diagonal PAPER baselines (instead of the 
horizontal E/W ones). Alternately, covariances could be estimated from all other baselines except the two being cross-multiplied 
when forming a power spectrum estimate. This method was used in \citet{parsons_et_al2014} (a similar method was also used in \citet{dillon_et_al2015}) in order to avoid suppressing the 
$21$ cm signal, and it's worth noting that the PAPER-32 results are likely less impacted from the issue of signal loss underestimation 
because of this very reason (however, they are affected by the error estimation issues described in Section \ref{sec:Error}, so 
we also regard those results as suspect and superseded by those of Kolopanis et al., \textit{submitted}).

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $144$ K, yielding $T_{sys} = 431$ K at $150$ MHz. 
This is lower than the $T_{sys}$ of $500$ K used in \citetalias{ali_et_al2015} because of several small miscalculation errors that were 
identified\footnote{For example, there was a missing a square root in going from a variance to a standard deviation.}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after cross-multiplying independent `even' and `odd' visibility measurements (which is done principally to avoid a noise bias). In \citetalias{ali_et_al2015}, a factor of $2$ was mistakenly used.
\item $N_{lst}$: The number of LST bins that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{lst} = 8$.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The updated limits in Kolopanis et al., \textit{submitted}
use three separation types.
\item $t_{int}$: Length of an independent integration of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time $t_{i}$
using the following:
\begin{equation}
t_{int} = t_{i} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{i}=43$ seconds, $t_{int}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{int} = 3857$ s. 
\item $N_{days}$: The total number of days of data analyzed. In \citetalias{ali_et_al2015}, this number was set to $135$. However, because we 
divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, 
because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a more realistic value of 
$N_{days}$ as:
\begin{equation}
 \frac{1}{N_{days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ is over LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ 
days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In \citetalias{ali_et_al2015}, this number was 
the total number of $30$ m East/West baselines used in the analysis. However, using the total number of baselines neglects 
the fact that baselines are averaged into groups (see Section \ref{sec:Boot}) for computational speed-up when cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{bls} = \frac{N_{bls\_total}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}},
\end{equation}
\noindent where, for the \citetalias{ali_et_al2015} analysis, $N_{gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{bls\_total}/N_{gps}$) and then as the square root of the number of cross-multiplied pairs ($\sqrt{N_{gps}^{2} - 
N_{gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$ instead 
of $51$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}

@@

As a consequence of our investigations, we have also revised power spectrum results from PAPER-64. In Kolopanis et al., (\textit{submitted}), we quote a best $2\sigma$ power spectrum upper limit of $(218.7$ mK$)^{2}$ at $k=0.3$ $h$Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK. The reasons for a previously underestimated limit and 
ways in which our new analysis differs can be summarized by the following:

\begin{itemize}
\item Signal loss, previously found to be $<2\%$ in \citetalias{ali_et_al2015}, was under-estimated by a factor of $\sim$$1000$ for inverse 
covariance weighting. For our new analysis, we use a regularized covariance weighting method to minimize loss (Section 
\ref{sec:Weight}). However, because our revised weighting method is not as aggressive as the former, our results are still a 
factor of $\sim$$10$ in mK higher than previous limits. Under-estimated signal loss therefore represents the bulk of our revision. We 
note that PAPER's analysis is not the first to under-estimate loss; results from the GMRT (\citealt{paciga_et_al2013}) were also 
revised from new signal loss calculations associated with their singular value decomposition foreground filter.
\item Power spectrum errors, originally computed by bootstrapping, were under-estimated for the data by a factor of $\sim$$2$ in mK due to oversampling data whose effective number of independent samples was reduced from fringe-rate filtering (Section \ref{sec:Boot}). 
\item Several factors used in an analytic expression to predict the noise-level in PAPER-64 data were revised, yielding a 
decrease in predicted sensitivity level by a factor of $\sim$$2$ in mK (Section \ref{sec:PSSense}). We have verified our revised 
prediction extensively using pure noise simulations. We note that our sensitivity prediction is revised by a factor less than our 
power spectrum result, implying that if taken at face value, the theoretical prediction for noise in \citetalias{ali_et_al2015} was too high for its data 
points.
\end{itemize}
